---
title: "ACTL4001 Capital Modelling and Risk Assessment"
author: "SL"
date: "2026-03-08"
output: html_document
---


3a). Aggregate loss simulation in R

1. Equipment Failure

```{r}
# =========================
# Q3(a) Equipment Failure
# Aggregate loss distribution
# =========================

set.seed(123)

n_sim <- 100000

# -------------------------
# 1. Inputs
# -------------------------
lambda_eq <- 0.1839667

mean_eq  <- 48000
disp_eq  <- 0.2312777
shape_eq <- 4.32
scale_eq <- 11100

# -------------------------
# 2. Simulate aggregate loss
# -------------------------
eq_loss <- numeric(n_sim)

for(i in 1:n_sim){
  
  n_claims <- rpois(1, lambda = lambda_eq)
  
  if(n_claims > 0){
    eq_loss[i] <- sum(
      rgamma(n_claims,
             shape = shape_eq,
             scale = scale_eq)
    )
  }
}

# -------------------------
# 3. Summary statistics
# -------------------------
var95_eq <- as.numeric(quantile(eq_loss, 0.95))
var99_eq <- as.numeric(quantile(eq_loss, 0.99))
var995_eq <- as.numeric(quantile(eq_loss, 0.995))

eq_summary <- data.frame(
  Metric = c("Expected annual loss",
             "Variance",
             "Standard deviation",
             "95% VaR",
             "99% VaR",
             "99.5% VaR"),
  Value = c(mean(eq_loss),
            var(eq_loss),
            sd(eq_loss),
            var95_eq,
            var99_eq,
            var995_eq)
)

eq_summary

# -------------------------
# 4. Plot aggregate loss distribution
# -------------------------
hist(eq_loss,
     breaks = 50,
     xlim = c(0, max(var99_eq * 1.15, max(eq_loss))),
     main = "Equipment Failure Aggregate Loss Distribution",
     xlab = "Annual Aggregate Loss",
     ylab = "Frequency",
     col = "lightgrey",
     border = "grey40")

abline(v = var95_eq, col = "blue", lty = 2, lwd = 2)
abline(v = var99_eq, col = "red",  lty = 2, lwd = 2)

legend("topright",
       legend = c("95% VaR", "99% VaR"),
       col = c("blue", "red"),
       lty = 2,
       lwd = 2,
       bty = "n")
```





3a). Aggregate loss simulation in R

2. Business Interruption

```{r}

# =========================
# Business Interruption Capital Modelling
# Q3(a) Aggregate Loss Distribution
# =========================

set.seed(123)

n_sim <- 100000

# -------------------------
# 1. Parameters
# -------------------------
# Frequency: Negative Binomial
r_bi <- 0.121553
p_bi <- 0.547205

# Severity: Gamma
shape_bi <- 0.3538
scale_bi <- 12327906.40

# -------------------------
# 2. Simulate aggregate losses
# -------------------------
# Step 1: simulate claim counts
n_claims <- rnbinom(n_sim, size = r_bi, prob = p_bi)

# Step 2: simulate aggregate loss for each year
bi_loss <- numeric(n_sim)

for (i in seq_len(n_sim)) {
  if (n_claims[i] > 0) {
    bi_loss[i] <- sum(rgamma(n_claims[i],
                             shape = shape_bi,
                             scale = scale_bi))
  }
}

# -------------------------
# 3. Summary statistics
# -------------------------
mean_loss <- mean(bi_loss)
var_loss  <- var(bi_loss)
sd_loss   <- sd(bi_loss)

VaR90  <- as.numeric(quantile(bi_loss, 0.90))
VaR95  <- as.numeric(quantile(bi_loss, 0.95))
VaR99  <- as.numeric(quantile(bi_loss, 0.99))
VaR995 <- as.numeric(quantile(bi_loss, 0.995))
VaR999 <- as.numeric(quantile(bi_loss, 0.999))

summary_table <- data.frame(
  Metric = c("Mean Annual Loss",
             "Variance",
             "Standard Deviation",
             "90th Percentile (VaR 90)",
             "95th Percentile (VaR 95)",
             "99th Percentile (VaR 99)",
             "99.5th Percentile (VaR 99.5)",
             "99.9th Percentile (VaR 99.9)"),
  Value = c(mean_loss, var_loss, sd_loss, VaR90, VaR95, VaR99, VaR995, VaR999)
)

print(summary_table)

# Optional: nicer display without scientific notation
summary_table$Value <- format(round(summary_table$Value, 2), big.mark = ",", scientific = FALSE)
print(summary_table)

# -------------------------
# 4. Plot 1: Full distribution
# -------------------------
hist(bi_loss,
     breaks = 100,
     main = "Business Interruption Aggregate Loss Distribution (Full Scale)",
     xlab = "Annual Aggregate Loss",
     col = "lightblue",
     border = "grey40")

abline(v = VaR95, col = "blue", lty = 2, lwd = 2)
abline(v = VaR99, col = "red", lty = 2, lwd = 2)

legend("topright",
       legend = c("95% VaR", "99% VaR"),
       col = c("blue", "red"),
       lty = 2,
       lwd = 2,
       bty = "n")

# -------------------------
# 5. Plot 2: Zoomed-in distribution
# -------------------------
hist(bi_loss,
     breaks = 100,
     xlim = c(0, VaR99 * 1.2),
     main = "Business Interruption Aggregate Loss Distribution (Zoomed)",
     xlab = "Annual Aggregate Loss",
     col = "lightblue",
     border = "grey40")

abline(v = VaR95, col = "blue", lty = 2, lwd = 2)
abline(v = VaR99, col = "red", lty = 2, lwd = 2)

legend("topright",
       legend = c("95% VaR", "99% VaR"),
       col = c("blue", "red"),
       lty = 2,
       lwd = 2,
       bty = "n")

# -------------------------
# 6. Additional diagnostics
# -------------------------
cat("Probability of zero aggregate loss =", mean(bi_loss == 0), "\n")
cat("Maximum simulated loss =", max(bi_loss), "\n")
```





3. Cargo Loss


```{r}
# =========================
# Cargo Loss Capital Modelling
# Q3(a) Aggregate loss distribution
# =========================

library(readxl)

cargo_data_raw <- read_excel("~/OneDrive - UNSW/Actuarial Institute/ACC1/SOA_2026_Case_Study_Materials/srcsc-2026-claims-cargo.xlsx")

set.seed(123)

n_sim <- 100000

# -------------------------
# 1. Inputs
# -------------------------
r_cargo <- 0.5370
p_cargo <- 0.6868

a_cargo <- 1.3832
b_cargo <- 14.3850

# Use observed cargo values from raw data
cargo_values <- cargo_data_raw$cargo_value
cargo_values <- cargo_values[!is.na(cargo_values) & cargo_values > 0]

# -------------------------
# 2. Simulate aggregate cargo loss
# -------------------------
cargo_loss <- numeric(n_sim)

for(i in 1:n_sim){
  
  n_claims <- rnbinom(1, size = r_cargo, prob = p_cargo)
  
  if(n_claims > 0){
    # Simulate proportional loss
    damage_ratio <- rbeta(n_claims, shape1 = a_cargo, shape2 = b_cargo)
    
    # Sample shipment values from observed portfolio
    sampled_values <- sample(cargo_values, size = n_claims, replace = TRUE)
    
    # Dollar loss = damage ratio × cargo value
    losses <- damage_ratio * sampled_values
    
    cargo_loss[i] <- sum(losses)
  }
}

# -------------------------
# 3. Summary table
# -------------------------
var95_cargo  <- as.numeric(quantile(cargo_loss, 0.95))
var99_cargo  <- as.numeric(quantile(cargo_loss, 0.99))
var995_cargo <- as.numeric(quantile(cargo_loss, 0.995))

cargo_summary <- data.frame(
  Metric = c("Expected annual loss",
             "Variance",
             "Standard deviation",
             "95% VaR",
             "99% VaR",
             "99.5% VaR"),
  Value = c(mean(cargo_loss),
            var(cargo_loss),
            sd(cargo_loss),
            var95_cargo,
            var99_cargo,
            var995_cargo)
)

cargo_summary


# -------------------------
# 4. Plotting
# -------------------------
hist(cargo_loss,
     breaks = 40,
     main = "Cargo Aggregate Loss Distribution",
     xlab = "Annual Aggregate Loss",
     col = "lightgrey",
     border = "black")

abline(v = var95_cargo, col = "blue", lwd = 2, lty = 2)
abline(v = var99_cargo, col = "red", lwd = 2, lty = 2)

legend("topright",
       legend = c("95% VaR", "99% VaR"),
       col = c("blue", "red"),
       lty = c(2, 2),
       lwd = 2,
       bty = "n")
```

4. Worker's Compensation



```{r}
# =========================
# Workers' Compensation Capital Modelling
# Q3(a) Aggregate Loss Distribution
# =========================

set.seed(123)

n_sim <- 100000

# -------------------------
# 1. Parameters
# -------------------------
# Frequency: Poisson
lambda_wc <- 0.017665

# Severity: Lognormal
mu_wc <- 7.619821
sigma2_wc <- 1.163305
sigma_wc <- sqrt(sigma2_wc)

# -------------------------
# 2. Simulate aggregate losses
# -------------------------
# Simulate annual claim counts
n_claims <- rpois(n_sim, lambda = lambda_wc)

# Simulate annual aggregate losses
wc_loss <- numeric(n_sim)

for (i in seq_len(n_sim)) {
  if (n_claims[i] > 0) {
    wc_loss[i] <- sum(rlnorm(n_claims[i],
                             meanlog = mu_wc,
                             sdlog = sigma_wc))
  }
}

# -------------------------
# 3. Summary statistics
# -------------------------
mean_loss <- mean(wc_loss)
var_loss  <- var(wc_loss)
sd_loss   <- sd(wc_loss)

VaR90  <- as.numeric(quantile(wc_loss, 0.90))
VaR95  <- as.numeric(quantile(wc_loss, 0.95))
VaR99  <- as.numeric(quantile(wc_loss, 0.99))
VaR995 <- as.numeric(quantile(wc_loss, 0.995))
VaR999 <- as.numeric(quantile(wc_loss, 0.999))

summary_table <- data.frame(
  Metric = c("Mean Annual Loss",
             "Variance",
             "Standard Deviation",
             "90th Percentile (VaR 90)",
             "95th Percentile (VaR 95)",
             "99th Percentile (VaR 99)",
             "99.5th Percentile (VaR 99.5)",
             "99.9th Percentile (VaR 99.9)"),
  Value = c(mean_loss, var_loss, sd_loss, VaR90, VaR95, VaR99, VaR995, VaR999)
)

print(summary_table)

# Optional formatting for report tables
summary_table_fmt <- summary_table
summary_table_fmt$Value <- format(round(summary_table_fmt$Value, 2),
                                  big.mark = ",",
                                  scientific = FALSE)
print(summary_table_fmt)

# -------------------------
# 4. Diagnostics
# -------------------------
prob_zero <- mean(wc_loss == 0)
max_loss  <- max(wc_loss)

cat("Probability of zero aggregate loss =", prob_zero, "\n")
cat("Maximum simulated loss =", max_loss, "\n")

# -------------------------
# 5. Plot 1: Full distribution
# -------------------------
hist(wc_loss,
     breaks = 100,
     main = "Workers' Compensation Aggregate Loss Distribution (Full Scale)",
     xlab = "Annual Aggregate Loss",
     col = "lightblue",
     border = "grey40")

abline(v = VaR95, col = "blue", lty = 2, lwd = 2)
abline(v = VaR99, col = "red", lty = 2, lwd = 2)

legend("topright",
       legend = c("95% VaR", "99% VaR"),
       col = c("blue", "red"),
       lty = 2,
       lwd = 2,
       bty = "n")

# -------------------------
# 6. Plot 2: Zoomed distribution
# -------------------------
hist(wc_loss,
     breaks = 100,
     xlim = c(0, VaR99 * 1.2),
     main = "Workers' Compensation Aggregate Loss Distribution (Zoomed)",
     xlab = "Annual Aggregate Loss",
     col = "lightblue",
     border = "grey40")

abline(v = VaR95, col = "blue", lty = 2, lwd = 2)
abline(v = VaR99, col = "red", lty = 2, lwd = 2)

legend("topright",
       legend = c("95% VaR", "99% VaR"),
       col = c("blue", "red"),
       lty = 2,
       lwd = 2,
       bty = "n")
```




3b).Stress testing for extreme scenarios (e.g., 1-in-100 year events)

Equipment Failure

```{r}
# =========================
# Equipment Failure Capital Modelling
# Q3(b) Stress testing for extreme scenarios
# =========================

rm(list = ls())
set.seed(123)

# -------------------------
# 1. Inputs
# -------------------------
n_sim <- 100000

# Baseline model from Q3(a)
lambda_eq <- 0.1839667
shape_eq  <- 4.32
scale_eq  <- 11100

# -------------------------
# 2. Scenario assumptions
# -------------------------
scenario_assumptions <- data.frame(
  Scenario = c("Baseline", "Moderate Stress", "Extreme Stress"),
  Lambda_Multiplier = c(1.00, 1.25, 1.50),
  Severity_Multiplier = c(1.00, 1.15, 1.30),
  Scenario_Description = c(
    "Normal operating conditions",
    "Elevated operational strain and minor system disruption",
    "Severe disruption such as solar storm or systemic malfunction"
  ),
  stringsAsFactors = FALSE
)

# -------------------------
# 3. Simulation function
# -------------------------
simulate_compound_poisson_gamma <- function(n_sim, lambda, shape, scale) {
  claim_count <- rpois(n_sim, lambda)
  agg_loss <- numeric(n_sim)
  
  for (i in seq_len(n_sim)) {
    if (claim_count[i] > 0) {
      agg_loss[i] <- sum(
        rgamma(claim_count[i], shape = shape, scale = scale)
      )
    }
  }
  
  list(
    agg_loss = agg_loss,
    claim_count = claim_count
  )
}

# -------------------------
# 4. Summary function
# -------------------------
summarise_scenario <- function(losses, counts, scenario_name) {
  data.frame(
    Scenario = scenario_name,
    Expected_Annual_Loss = mean(losses),
    Variance = var(losses),
    Standard_Deviation = sd(losses),
    VaR_95 = as.numeric(quantile(losses, 0.95)),
    VaR_99 = as.numeric(quantile(losses, 0.99)),
    VaR_995 = as.numeric(quantile(losses, 0.995)),
    Prob_Zero_Loss = mean(losses == 0),
    Prob_1_Claim = mean(counts == 1),
    Prob_2_Claims = mean(counts == 2),
    Prob_3plus_Claims = mean(counts >= 3)
  )
}

# -------------------------
# 5. Run all scenarios
# -------------------------
scenario_results <- list()
summary_table <- data.frame()

for (i in seq_len(nrow(scenario_assumptions))) {
  
  scenario_name <- scenario_assumptions$Scenario[i]
  lambda_i <- lambda_eq * scenario_assumptions$Lambda_Multiplier[i]
  scale_i  <- scale_eq  * scenario_assumptions$Severity_Multiplier[i]
  
  sim_i <- simulate_compound_poisson_gamma(
    n_sim = n_sim,
    lambda = lambda_i,
    shape = shape_eq,
    scale = scale_i
  )
  
  scenario_results[[scenario_name]] <- sim_i
  
  summary_i <- summarise_scenario(
    losses = sim_i$agg_loss,
    counts = sim_i$claim_count,
    scenario_name = scenario_name
  )
  
  summary_table <- rbind(summary_table, summary_i)
}

# -------------------------
# 6. Add stress ratios vs baseline
# -------------------------
baseline_expected <- summary_table$Expected_Annual_Loss[summary_table$Scenario == "Baseline"]
baseline_var99    <- summary_table$VaR_99[summary_table$Scenario == "Baseline"]
baseline_var995   <- summary_table$VaR_995[summary_table$Scenario == "Baseline"]

summary_table$Expected_Loss_Ratio_vs_Baseline <- summary_table$Expected_Annual_Loss / baseline_expected
summary_table$VaR99_Ratio_vs_Baseline         <- summary_table$VaR_99 / baseline_var99
summary_table$VaR995_Ratio_vs_Baseline        <- summary_table$VaR_995 / baseline_var995

# -------------------------
# 7. Main report-style comparison table
# -------------------------
main_scenario_table <- data.frame(
  Metric = c(
    "Expected annual loss",
    "Variance",
    "Standard deviation",
    "95% VaR",
    "99% VaR",
    "99.5% VaR",
    "Expected loss ratio vs baseline",
    "99% VaR ratio vs baseline",
    "99.5% VaR ratio vs baseline"
  ),
  
  Baseline = c(
    round(summary_table$Expected_Annual_Loss[summary_table$Scenario == "Baseline"], 0),
    round(summary_table$Variance[summary_table$Scenario == "Baseline"], 0),
    round(summary_table$Standard_Deviation[summary_table$Scenario == "Baseline"], 0),
    round(summary_table$VaR_95[summary_table$Scenario == "Baseline"], 0),
    round(summary_table$VaR_99[summary_table$Scenario == "Baseline"], 0),
    round(summary_table$VaR_995[summary_table$Scenario == "Baseline"], 0),
    round(summary_table$Expected_Loss_Ratio_vs_Baseline[summary_table$Scenario == "Baseline"], 2),
    round(summary_table$VaR99_Ratio_vs_Baseline[summary_table$Scenario == "Baseline"], 2),
    round(summary_table$VaR995_Ratio_vs_Baseline[summary_table$Scenario == "Baseline"], 2)
  ),
  
  Moderate_Stress = c(
    round(summary_table$Expected_Annual_Loss[summary_table$Scenario == "Moderate Stress"], 0),
    round(summary_table$Variance[summary_table$Scenario == "Moderate Stress"], 0),
    round(summary_table$Standard_Deviation[summary_table$Scenario == "Moderate Stress"], 0),
    round(summary_table$VaR_95[summary_table$Scenario == "Moderate Stress"], 0),
    round(summary_table$VaR_99[summary_table$Scenario == "Moderate Stress"], 0),
    round(summary_table$VaR_995[summary_table$Scenario == "Moderate Stress"], 0),
    round(summary_table$Expected_Loss_Ratio_vs_Baseline[summary_table$Scenario == "Moderate Stress"], 2),
    round(summary_table$VaR99_Ratio_vs_Baseline[summary_table$Scenario == "Moderate Stress"], 2),
    round(summary_table$VaR995_Ratio_vs_Baseline[summary_table$Scenario == "Moderate Stress"], 2)
  ),
  
  Extreme_Stress = c(
    round(summary_table$Expected_Annual_Loss[summary_table$Scenario == "Extreme Stress"], 0),
    round(summary_table$Variance[summary_table$Scenario == "Extreme Stress"], 0),
    round(summary_table$Standard_Deviation[summary_table$Scenario == "Extreme Stress"], 0),
    round(summary_table$VaR_95[summary_table$Scenario == "Extreme Stress"], 0),
    round(summary_table$VaR_99[summary_table$Scenario == "Extreme Stress"], 0),
    round(summary_table$VaR_995[summary_table$Scenario == "Extreme Stress"], 0),
    round(summary_table$Expected_Loss_Ratio_vs_Baseline[summary_table$Scenario == "Extreme Stress"], 2),
    round(summary_table$VaR99_Ratio_vs_Baseline[summary_table$Scenario == "Extreme Stress"], 2),
    round(summary_table$VaR995_Ratio_vs_Baseline[summary_table$Scenario == "Extreme Stress"], 2)
  )
)

cat("=== Equipment Failure: Stress Testing Comparison Table ===\n")
print(main_scenario_table)

# Optional formatted version for Word-ready copying
main_scenario_table_fmt <- main_scenario_table
main_scenario_table_fmt[-1] <- lapply(main_scenario_table_fmt[-1], function(x) {
  ifelse(x >= 1000, format(x, big.mark = ",", scientific = FALSE), as.character(x))
})
cat("\n=== Formatted Equipment Failure Stress Table ===\n")
print(main_scenario_table_fmt, row.names = FALSE)

# -------------------------
# 8. Appendix-style diagnostic table
# -------------------------
diagnostic_table <- data.frame(
  Metric = c(
    "Probability of zero loss",
    "Probability of 1 claim",
    "Probability of 2 claims",
    "Probability of 3+ claims"
  ),
  
  Baseline = c(
    round(summary_table$Prob_Zero_Loss[summary_table$Scenario == "Baseline"], 4),
    round(summary_table$Prob_1_Claim[summary_table$Scenario == "Baseline"], 4),
    round(summary_table$Prob_2_Claims[summary_table$Scenario == "Baseline"], 4),
    round(summary_table$Prob_3plus_Claims[summary_table$Scenario == "Baseline"], 4)
  ),
  
  Moderate_Stress = c(
    round(summary_table$Prob_Zero_Loss[summary_table$Scenario == "Moderate Stress"], 4),
    round(summary_table$Prob_1_Claim[summary_table$Scenario == "Moderate Stress"], 4),
    round(summary_table$Prob_2_Claims[summary_table$Scenario == "Moderate Stress"], 4),
    round(summary_table$Prob_3plus_Claims[summary_table$Scenario == "Moderate Stress"], 4)
  ),
  
  Extreme_Stress = c(
    round(summary_table$Prob_Zero_Loss[summary_table$Scenario == "Extreme Stress"], 4),
    round(summary_table$Prob_1_Claim[summary_table$Scenario == "Extreme Stress"], 4),
    round(summary_table$Prob_2_Claims[summary_table$Scenario == "Extreme Stress"], 4),
    round(summary_table$Prob_3plus_Claims[summary_table$Scenario == "Extreme Stress"], 4)
  )
)

cat("\n=== Equipment Failure: Diagnostic Table ===\n")
print(diagnostic_table)

# -------------------------
# 9. Plot baseline distribution with VaR lines
# -------------------------
eq_loss_baseline <- scenario_results[["Baseline"]]$agg_loss
VaR95_eq  <- as.numeric(quantile(eq_loss_baseline, 0.95))
VaR99_eq  <- as.numeric(quantile(eq_loss_baseline, 0.99))
VaR995_eq <- as.numeric(quantile(eq_loss_baseline, 0.995))

hist(eq_loss_baseline,
     breaks = 40,
     xlim = c(0, max(VaR995_eq * 1.2, max(eq_loss_baseline))),
     main = "Equipment Failure Aggregate Loss Distribution (Baseline)",
     xlab = "Annual Aggregate Loss",
     ylab = "Frequency",
     col = "lightgrey",
     border = "black")

abline(v = VaR95_eq,  col = "blue",  lwd = 2, lty = 2)
abline(v = VaR99_eq,  col = "red",   lwd = 2, lty = 2)
abline(v = VaR995_eq, col = "darkgreen", lwd = 2, lty = 2)

legend("topright",
       legend = c("95% VaR", "99% VaR", "99.5% VaR"),
       col = c("blue", "red", "darkgreen"),
       lty = c(2, 2, 2),
       lwd = 2,
       bty = "n")

# -------------------------
# 10. Plot scenario comparison
# -------------------------
hist(scenario_results[["Baseline"]]$agg_loss,
     breaks = 40,
     xlim = c(0, max(
       quantile(scenario_results[["Extreme Stress"]]$agg_loss, 0.995) * 1.2,
       max(scenario_results[["Extreme Stress"]]$agg_loss)
     )),
     main = "Equipment Failure: Scenario Comparison",
     xlab = "Annual Aggregate Loss",
     ylab = "Frequency",
     col = rgb(0, 0, 1, 0.25),
     border = "blue")

hist(scenario_results[["Moderate Stress"]]$agg_loss,
     breaks = 40,
     add = TRUE,
     col = rgb(1, 0.5, 0, 0.25),
     border = "orange")

hist(scenario_results[["Extreme Stress"]]$agg_loss,
     breaks = 40,
     add = TRUE,
     col = rgb(1, 0, 0, 0.25),
     border = "red")

legend("topright",
       legend = c("Baseline", "Moderate Stress", "Extreme Stress"),
       fill = c(
         rgb(0, 0, 1, 0.25),
         rgb(1, 0.5, 0, 0.25),
         rgb(1, 0, 0, 0.25)
       ),
       border = c("blue", "orange", "red"),
       bty = "n")
```


3b).Stress testing for extreme scenarios (e.g., 1-in-100 year events)

Business Interruption

```{r}
# =========================
# Business Interruption Capital Modelling
# Q3(b) Stress testing for extreme scenarios
# =========================

rm(list = ls())
set.seed(123)

# -------------------------
# 1. Inputs
# -------------------------
n_sim <- 100000

# Baseline model from Q3(a)
r_bi <- 0.121553
p_bi <- 0.547205
shape_bi <- 0.3538
scale_bi <- 12327906.40

# -------------------------
# 2. Scenario assumptions
# -------------------------
scenario_assumptions <- data.frame(
  Scenario = c("Baseline", "Moderate Stress", "Extreme Stress"),
  Frequency_Multiplier = c(1.00, 1.20, 1.40),
  Severity_Multiplier = c(1.00, 1.25, 1.50),
  stringsAsFactors = FALSE
)

# -------------------------
# 3. Simulation function
# -------------------------
simulate_compound_nbinom_gamma <- function(n_sim, size, prob, freq_mult, shape, scale, sev_mult) {
  claim_count <- rnbinom(n_sim, size = size * freq_mult, prob = prob)
  agg_loss <- numeric(n_sim)
  
  for (i in seq_len(n_sim)) {
    if (claim_count[i] > 0) {
      agg_loss[i] <- sum(
        rgamma(claim_count[i], shape = shape, scale = scale * sev_mult)
      )
    }
  }
  
  list(agg_loss = agg_loss, claim_count = claim_count)
}

# -------------------------
# 4. Summary function
# -------------------------
summarise_scenario <- function(losses, counts, scenario_name) {
  data.frame(
    Scenario = scenario_name,
    Expected_Annual_Loss = mean(losses),
    Variance = var(losses),
    Standard_Deviation = sd(losses),
    VaR_95 = as.numeric(quantile(losses, 0.95)),
    VaR_99 = as.numeric(quantile(losses, 0.99)),
    VaR_995 = as.numeric(quantile(losses, 0.995)),
    Prob_Zero_Loss = mean(losses == 0),
    Prob_1_Claim = mean(counts == 1),
    Prob_2plus_Claims = mean(counts >= 2)
  )
}

# -------------------------
# 5. Run scenarios
# -------------------------
scenario_results <- list()
summary_table <- data.frame()

for (i in seq_len(nrow(scenario_assumptions))) {
  sim_i <- simulate_compound_nbinom_gamma(
    n_sim = n_sim,
    size = r_bi,
    prob = p_bi,
    freq_mult = scenario_assumptions$Frequency_Multiplier[i],
    shape = shape_bi,
    scale = scale_bi,
    sev_mult = scenario_assumptions$Severity_Multiplier[i]
  )
  
  scenario_results[[scenario_assumptions$Scenario[i]]] <- sim_i
  
  summary_i <- summarise_scenario(
    losses = sim_i$agg_loss,
    counts = sim_i$claim_count,
    scenario_name = scenario_assumptions$Scenario[i]
  )
  
  summary_table <- rbind(summary_table, summary_i)
}

# -------------------------
# 6. Add ratios vs baseline
# -------------------------
baseline_expected <- summary_table$Expected_Annual_Loss[summary_table$Scenario == "Baseline"]
baseline_var99 <- summary_table$VaR_99[summary_table$Scenario == "Baseline"]
baseline_var995 <- summary_table$VaR_995[summary_table$Scenario == "Baseline"]

summary_table$Expected_Loss_Ratio_vs_Baseline <- summary_table$Expected_Annual_Loss / baseline_expected
summary_table$VaR99_Ratio_vs_Baseline <- summary_table$VaR_99 / baseline_var99
summary_table$VaR995_Ratio_vs_Baseline <- summary_table$VaR_995 / baseline_var995

# -------------------------
# 7. Main report-style table
# -------------------------
main_scenario_table <- data.frame(
  Metric = c("Expected annual loss", "Variance", "Standard deviation",
             "95% VaR", "99% VaR", "99.5% VaR",
             "Expected loss ratio vs baseline",
             "99% VaR ratio vs baseline",
             "99.5% VaR ratio vs baseline"),
  Baseline = c(
    round(summary_table$Expected_Annual_Loss[1], 0),
    round(summary_table$Variance[1], 0),
    round(summary_table$Standard_Deviation[1], 0),
    round(summary_table$VaR_95[1], 0),
    round(summary_table$VaR_99[1], 0),
    round(summary_table$VaR_995[1], 0),
    round(summary_table$Expected_Loss_Ratio_vs_Baseline[1], 2),
    round(summary_table$VaR99_Ratio_vs_Baseline[1], 2),
    round(summary_table$VaR995_Ratio_vs_Baseline[1], 2)
  ),
  Moderate_Stress = c(
    round(summary_table$Expected_Annual_Loss[2], 0),
    round(summary_table$Variance[2], 0),
    round(summary_table$Standard_Deviation[2], 0),
    round(summary_table$VaR_95[2], 0),
    round(summary_table$VaR_99[2], 0),
    round(summary_table$VaR_995[2], 0),
    round(summary_table$Expected_Loss_Ratio_vs_Baseline[2], 2),
    round(summary_table$VaR99_Ratio_vs_Baseline[2], 2),
    round(summary_table$VaR995_Ratio_vs_Baseline[2], 2)
  ),
  Extreme_Stress = c(
    round(summary_table$Expected_Annual_Loss[3], 0),
    round(summary_table$Variance[3], 0),
    round(summary_table$Standard_Deviation[3], 0),
    round(summary_table$VaR_95[3], 0),
    round(summary_table$VaR_99[3], 0),
    round(summary_table$VaR_995[3], 0),
    round(summary_table$Expected_Loss_Ratio_vs_Baseline[3], 2),
    round(summary_table$VaR99_Ratio_vs_Baseline[3], 2),
    round(summary_table$VaR995_Ratio_vs_Baseline[3], 2)
  )
)

print(main_scenario_table)

# -------------------------
# 8. Diagnostic table
# -------------------------
diagnostic_table <- data.frame(
  Metric = c("Probability of zero loss", "Probability of 1 claim", "Probability of 2+ claims"),
  Baseline = c(round(summary_table$Prob_Zero_Loss[1], 4),
               round(summary_table$Prob_1_Claim[1], 4),
               round(summary_table$Prob_2plus_Claims[1], 4)),
  Moderate_Stress = c(round(summary_table$Prob_Zero_Loss[2], 4),
                      round(summary_table$Prob_1_Claim[2], 4),
                      round(summary_table$Prob_2plus_Claims[2], 4)),
  Extreme_Stress = c(round(summary_table$Prob_Zero_Loss[3], 4),
                     round(summary_table$Prob_1_Claim[3], 4),
                     round(summary_table$Prob_2plus_Claims[3], 4))
)

print(diagnostic_table)

# -------------------------
# 9. Plots
# -------------------------
hist(scenario_results[["Baseline"]]$agg_loss,
     breaks = 40,
     main = "Business Interruption Aggregate Loss Distribution (Baseline)",
     xlab = "Annual Aggregate Loss",
     col = "lightgrey",
     border = "black")

hist(scenario_results[["Baseline"]]$agg_loss,
     breaks = 40,
     main = "Business Interruption: Scenario Comparison",
     xlab = "Annual Aggregate Loss",
     col = rgb(0, 0, 1, 0.25),
     border = "blue")

hist(scenario_results[["Moderate Stress"]]$agg_loss,
     breaks = 40, add = TRUE,
     col = rgb(1, 0.5, 0, 0.25),
     border = "orange")

hist(scenario_results[["Extreme Stress"]]$agg_loss,
     breaks = 40, add = TRUE,
     col = rgb(1, 0, 0, 0.25),
     border = "red")

legend("topright",
       legend = c("Baseline", "Moderate Stress", "Extreme Stress"),
       fill = c(rgb(0, 0, 1, 0.25), rgb(1, 0.5, 0, 0.25), rgb(1, 0, 0, 0.25)),
       border = c("blue", "orange", "red"),
       bty = "n")
```


3b).Stress testing for extreme scenarios (e.g., 1-in-100 year events)

Cargo Loss

```{r}
# =========================
# Cargo Loss Capital Modelling
# Q3(b) Stress testing for extreme scenarios
# =========================

rm(list = ls())
set.seed(123)

library(readxl)

cargo_data_raw <- read_excel("~/OneDrive - UNSW/Actuarial Institute/ACC1/SOA_2026_Case_Study_Materials/srcsc-2026-claims-cargo.xlsx")

# -------------------------
# 1. Inputs
# -------------------------
n_sim <- 100000

r_cargo <- 0.5370
p_cargo <- 0.6868
a_cargo <- 1.3832
b_cargo <- 14.3850

cargo_values <- cargo_data_raw$cargo_value
cargo_values <- cargo_values[!is.na(cargo_values) & cargo_values > 0]

# -------------------------
# 2. Scenario assumptions
# -------------------------
scenario_assumptions <- data.frame(
  Scenario = c("Baseline", "Moderate Stress", "Extreme Stress"),
  Frequency_Multiplier = c(1.00, 1.20, 1.40),
  Severity_Multiplier = c(1.00, 1.15, 1.30),
  stringsAsFactors = FALSE
)

# -------------------------
# 3. Simulation function
# -------------------------
simulate_compound_nbinom_beta_cargo <- function(n_sim, size, prob, freq_mult, alpha, beta, sev_mult, cargo_values) {
  claim_count <- rnbinom(n_sim, size = size * freq_mult, prob = prob)
  agg_loss <- numeric(n_sim)
  
  for (i in seq_len(n_sim)) {
    if (claim_count[i] > 0) {
      damage_ratio <- rbeta(claim_count[i], shape1 = alpha, shape2 = beta)
      sampled_values <- sample(cargo_values, size = claim_count[i], replace = TRUE)
      losses <- damage_ratio * sampled_values * sev_mult
      agg_loss[i] <- sum(losses)
    }
  }
  
  list(agg_loss = agg_loss, claim_count = claim_count)
}

# -------------------------
# 4. Summary function
# -------------------------
summarise_scenario <- function(losses, counts, scenario_name) {
  data.frame(
    Scenario = scenario_name,
    Expected_Annual_Loss = mean(losses),
    Variance = var(losses),
    Standard_Deviation = sd(losses),
    VaR_95 = as.numeric(quantile(losses, 0.95)),
    VaR_99 = as.numeric(quantile(losses, 0.99)),
    VaR_995 = as.numeric(quantile(losses, 0.995)),
    Prob_Zero_Loss = mean(losses == 0),
    Prob_1_Claim = mean(counts == 1),
    Prob_2_Claims = mean(counts == 2),
    Prob_3plus_Claims = mean(counts >= 3)
  )
}

# -------------------------
# 5. Run scenarios
# -------------------------
scenario_results <- list()
summary_table <- data.frame()

for (i in seq_len(nrow(scenario_assumptions))) {
  sim_i <- simulate_compound_nbinom_beta_cargo(
    n_sim = n_sim,
    size = r_cargo,
    prob = p_cargo,
    freq_mult = scenario_assumptions$Frequency_Multiplier[i],
    alpha = a_cargo,
    beta = b_cargo,
    sev_mult = scenario_assumptions$Severity_Multiplier[i],
    cargo_values = cargo_values
  )
  
  scenario_results[[scenario_assumptions$Scenario[i]]] <- sim_i
  
  summary_i <- summarise_scenario(
    losses = sim_i$agg_loss,
    counts = sim_i$claim_count,
    scenario_name = scenario_assumptions$Scenario[i]
  )
  
  summary_table <- rbind(summary_table, summary_i)
}

# -------------------------
# 6. Add ratios vs baseline
# -------------------------
baseline_expected <- summary_table$Expected_Annual_Loss[1]
baseline_var99 <- summary_table$VaR_99[1]
baseline_var995 <- summary_table$VaR_995[1]

summary_table$Expected_Loss_Ratio_vs_Baseline <- summary_table$Expected_Annual_Loss / baseline_expected
summary_table$VaR99_Ratio_vs_Baseline <- summary_table$VaR_99 / baseline_var99
summary_table$VaR995_Ratio_vs_Baseline <- summary_table$VaR_995 / baseline_var995

# -------------------------
# 7. Main table
# -------------------------
main_scenario_table <- data.frame(
  Metric = c("Expected annual loss", "Variance", "Standard deviation",
             "95% VaR", "99% VaR", "99.5% VaR",
             "Expected loss ratio vs baseline",
             "99% VaR ratio vs baseline",
             "99.5% VaR ratio vs baseline"),
  Baseline = c(round(summary_table$Expected_Annual_Loss[1], 0),
               round(summary_table$Variance[1], 0),
               round(summary_table$Standard_Deviation[1], 0),
               round(summary_table$VaR_95[1], 0),
               round(summary_table$VaR_99[1], 0),
               round(summary_table$VaR_995[1], 0),
               round(summary_table$Expected_Loss_Ratio_vs_Baseline[1], 2),
               round(summary_table$VaR99_Ratio_vs_Baseline[1], 2),
               round(summary_table$VaR995_Ratio_vs_Baseline[1], 2)),
  Moderate_Stress = c(round(summary_table$Expected_Annual_Loss[2], 0),
                      round(summary_table$Variance[2], 0),
                      round(summary_table$Standard_Deviation[2], 0),
                      round(summary_table$VaR_95[2], 0),
                      round(summary_table$VaR_99[2], 0),
                      round(summary_table$VaR_995[2], 0),
                      round(summary_table$Expected_Loss_Ratio_vs_Baseline[2], 2),
                      round(summary_table$VaR99_Ratio_vs_Baseline[2], 2),
                      round(summary_table$VaR995_Ratio_vs_Baseline[2], 2)),
  Extreme_Stress = c(round(summary_table$Expected_Annual_Loss[3], 0),
                     round(summary_table$Variance[3], 0),
                     round(summary_table$Standard_Deviation[3], 0),
                     round(summary_table$VaR_95[3], 0),
                     round(summary_table$VaR_99[3], 0),
                     round(summary_table$VaR_995[3], 0),
                     round(summary_table$Expected_Loss_Ratio_vs_Baseline[3], 2),
                     round(summary_table$VaR99_Ratio_vs_Baseline[3], 2),
                     round(summary_table$VaR995_Ratio_vs_Baseline[3], 2))
)

print(main_scenario_table)

# -------------------------
# 8. Plots
# -------------------------
hist(scenario_results[["Baseline"]]$agg_loss,
     breaks = 40,
     main = "Cargo Loss Aggregate Loss Distribution (Baseline)",
     xlab = "Annual Aggregate Loss",
     col = "lightgrey",
     border = "black")

hist(scenario_results[["Baseline"]]$agg_loss,
     breaks = 40,
     main = "Cargo Loss: Scenario Comparison",
     xlab = "Annual Aggregate Loss",
     col = rgb(0, 0, 1, 0.25),
     border = "blue")

hist(scenario_results[["Moderate Stress"]]$agg_loss,
     breaks = 40, add = TRUE,
     col = rgb(1, 0.5, 0, 0.25),
     border = "orange")

hist(scenario_results[["Extreme Stress"]]$agg_loss,
     breaks = 40, add = TRUE,
     col = rgb(1, 0, 0, 0.25),
     border = "red")

legend("topright",
       legend = c("Baseline", "Moderate Stress", "Extreme Stress"),
       fill = c(rgb(0, 0, 1, 0.25), rgb(1, 0.5, 0, 0.25), rgb(1, 0, 0, 0.25)),
       border = c("blue", "orange", "red"),
       bty = "n")
```



```{r}
# =========================
# Workers' Compensation Capital Modelling
# Q3(b) Stress testing for extreme scenarios
# =========================

rm(list = ls())
set.seed(123)

# -------------------------
# 1. Inputs
# -------------------------
n_sim <- 100000

lambda_wc <- 0.017665
mu_wc <- 7.619821
sigma2_wc <- 1.163305
sigma_wc <- sqrt(sigma2_wc)

# -------------------------
# 2. Scenario assumptions
# -------------------------
scenario_assumptions <- data.frame(
  Scenario = c("Baseline", "Moderate Stress", "Extreme Stress"),
  Frequency_Multiplier = c(1.00, 1.20, 1.40),
  Severity_Shift = c(0.00, 0.10, 0.20),
  stringsAsFactors = FALSE
)

# -------------------------
# 3. Simulation function
# -------------------------
simulate_compound_poisson_lognormal <- function(n_sim, lambda, freq_mult, mu, sigma, mu_shift) {
  claim_count <- rpois(n_sim, lambda = lambda * freq_mult)
  agg_loss <- numeric(n_sim)
  
  for (i in seq_len(n_sim)) {
    if (claim_count[i] > 0) {
      agg_loss[i] <- sum(
        rlnorm(claim_count[i], meanlog = mu + mu_shift, sdlog = sigma)
      )
    }
  }
  
  list(agg_loss = agg_loss, claim_count = claim_count)
}

# -------------------------
# 4. Summary function
# -------------------------
summarise_scenario <- function(losses, counts, scenario_name) {
  data.frame(
    Scenario = scenario_name,
    Expected_Annual_Loss = mean(losses),
    Variance = var(losses),
    Standard_Deviation = sd(losses),
    VaR_95 = as.numeric(quantile(losses, 0.95)),
    VaR_99 = as.numeric(quantile(losses, 0.99)),
    VaR_995 = as.numeric(quantile(losses, 0.995)),
    Prob_Zero_Loss = mean(losses == 0),
    Prob_1_Claim = mean(counts == 1),
    Prob_2plus_Claims = mean(counts >= 2)
  )
}

# -------------------------
# 5. Run scenarios
# -------------------------
scenario_results <- list()
summary_table <- data.frame()

for (i in seq_len(nrow(scenario_assumptions))) {
  sim_i <- simulate_compound_poisson_lognormal(
    n_sim = n_sim,
    lambda = lambda_wc,
    freq_mult = scenario_assumptions$Frequency_Multiplier[i],
    mu = mu_wc,
    sigma = sigma_wc,
    mu_shift = scenario_assumptions$Severity_Shift[i]
  )
  
  scenario_results[[scenario_assumptions$Scenario[i]]] <- sim_i
  
  summary_i <- summarise_scenario(
    losses = sim_i$agg_loss,
    counts = sim_i$claim_count,
    scenario_name = scenario_assumptions$Scenario[i]
  )
  
  summary_table <- rbind(summary_table, summary_i)
}

# -------------------------
# 6. Add ratios vs baseline
# -------------------------
baseline_expected <- summary_table$Expected_Annual_Loss[1]
baseline_var99 <- summary_table$VaR_99[1]
baseline_var995 <- summary_table$VaR_995[1]

summary_table$Expected_Loss_Ratio_vs_Baseline <- summary_table$Expected_Annual_Loss / baseline_expected
summary_table$VaR99_Ratio_vs_Baseline <- summary_table$VaR_99 / baseline_var99
summary_table$VaR995_Ratio_vs_Baseline <- summary_table$VaR_995 / baseline_var995

# -------------------------
# 7. Main table
# -------------------------
main_scenario_table <- data.frame(
  Metric = c("Expected annual loss", "Variance", "Standard deviation",
             "95% VaR", "99% VaR", "99.5% VaR",
             "Expected loss ratio vs baseline",
             "99% VaR ratio vs baseline",
             "99.5% VaR ratio vs baseline"),
  Baseline = c(round(summary_table$Expected_Annual_Loss[1], 0),
               round(summary_table$Variance[1], 0),
               round(summary_table$Standard_Deviation[1], 0),
               round(summary_table$VaR_95[1], 0),
               round(summary_table$VaR_99[1], 0),
               round(summary_table$VaR_995[1], 0),
               round(summary_table$Expected_Loss_Ratio_vs_Baseline[1], 2),
               round(summary_table$VaR99_Ratio_vs_Baseline[1], 2),
               round(summary_table$VaR995_Ratio_vs_Baseline[1], 2)),
  Moderate_Stress = c(round(summary_table$Expected_Annual_Loss[2], 0),
                      round(summary_table$Variance[2], 0),
                      round(summary_table$Standard_Deviation[2], 0),
                      round(summary_table$VaR_95[2], 0),
                      round(summary_table$VaR_99[2], 0),
                      round(summary_table$VaR_995[2], 0),
                      round(summary_table$Expected_Loss_Ratio_vs_Baseline[2], 2),
                      round(summary_table$VaR99_Ratio_vs_Baseline[2], 2),
                      round(summary_table$VaR995_Ratio_vs_Baseline[2], 2)),
  Extreme_Stress = c(round(summary_table$Expected_Annual_Loss[3], 0),
                     round(summary_table$Variance[3], 0),
                     round(summary_table$Standard_Deviation[3], 0),
                     round(summary_table$VaR_95[3], 0),
                     round(summary_table$VaR_99[3], 0),
                     round(summary_table$VaR_995[3], 0),
                     round(summary_table$Expected_Loss_Ratio_vs_Baseline[3], 2),
                     round(summary_table$VaR99_Ratio_vs_Baseline[3], 2),
                     round(summary_table$VaR995_Ratio_vs_Baseline[3], 2))
)

print(main_scenario_table)

# -------------------------
# 8. Plots
# -------------------------
hist(scenario_results[["Baseline"]]$agg_loss,
     breaks = 40,
     main = "Workers' Compensation Aggregate Loss Distribution (Baseline)",
     xlab = "Annual Aggregate Loss",
     col = "lightgrey",
     border = "black")

hist(scenario_results[["Baseline"]]$agg_loss,
     breaks = 40,
     main = "Workers' Compensation: Scenario Comparison",
     xlab = "Annual Aggregate Loss",
     col = rgb(0, 0, 1, 0.25),
     border = "blue")

hist(scenario_results[["Moderate Stress"]]$agg_loss,
     breaks = 40, add = TRUE,
     col = rgb(1, 0.5, 0, 0.25),
     border = "orange")

hist(scenario_results[["Extreme Stress"]]$agg_loss,
     breaks = 40, add = TRUE,
     col = rgb(1, 0, 0, 0.25),
     border = "red")

legend("topright",
       legend = c("Baseline", "Moderate Stress", "Extreme Stress"),
       fill = c(rgb(0, 0, 1, 0.25), rgb(1, 0.5, 0, 0.25), rgb(1, 0, 0, 0.25)),
       border = c("blue", "orange", "red"),
       bty = "n")
```


```{r}
# prepare data
library(dplyr)
library(ggplot2)

loss_data <- data.frame(
  Loss = c(equipment_loss, cargo_loss, bi_loss, workers_loss),
  Hazard = factor(rep(c("Equipment Failure",
                        "Cargo Loss",
                        "Business Interruption",
                        "Workers Compensation"),
                      each = n_sim))
)


# Professional density plot
ggplot(loss_data, aes(x = Loss, fill = Hazard)) +
  
  geom_density(alpha = 0.5, color = NA) +
  
  facet_wrap(~Hazard, scales = "free") +
  
  scale_fill_manual(values = c(
    "#2E86AB",
    "#E67E22",
    "#2ECC71",
    "#8E44AD"
  )) +
  
  labs(
    title = "Simulated Aggregate Annual Loss Distributions",
    subtitle = "Monte Carlo simulation (50,000 years)",
    x = "Annual Aggregate Loss ($)",
    y = "Density"
  ) +
  
  theme_minimal(base_size = 12) +
  
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    strip.text = element_text(face = "bold")
  )

#Export for the report
ggsave("aggregate_loss_distribution.png",
       width = 8,
       height = 5,
       dpi = 300)

```












